### 2015-10-01 Manuscript/carkeek_grid.tex
first commit

%--------------------------------------||--------------------------------------%

\documentclass[11pt,letterpaper]{article} %  font size

%-----------------------------------PACKAGES-----------------------------------%

\usepackage[T1]{fontenc} % Choose an output font encoding (T1) that has support for the accented characters used by the most widespread European languages
\usepackage[utf8]{inputenc} % Allow input of accented characters (and more...)
\usepackage{graphicx} %  figures
\usepackage[round]{natbib} % names in citations
\usepackage{lineno} % line numbers
\usepackage{authblk} % allows more intuitive formatting for multiple authors/affiliations
\usepackage[margin=1in]{geometry} % make the margins 1 inch on all sides of the document
%\usepackage{amsmath} % useful for formatting math stuff, especially complex equations
%\usepackage{pdflscape} % rotate table into landscape mode
%\usepackage{subfigure} % side by side figures
%\usepackage{longtable} % for tables that span multiple pages.
\usepackage{setspace} % double spacing

%----------------------------------FORMATTING-----------------------------------

%\topmargin -1cm %0.0cm
%\textwidth 16cm % what does this do?
%\textheight 21cm % what does this do?
%\footskip 1.0cm % what does this do?
%\oddsidemargin 0.0cm

\date{\today}
\doublespacing % initiate double spacing (package setspace)
%\linespread{2} % alternate method of double spacing

%------------------------------------TITLE--------------------------------------

\title{There once was a grid at ol' Carkeek}

%-----------------------------------AUTHORS-------------------------------------

% using package 'authblk':
\author[1]{First Author\thanks{first.author@funstuff.com}}

\author[1,2]{Second Author}

\author[2]{Third Author}

\affil[1]{Department of Computer Science, \LaTeX\ University}
\affil[2]{Department of Mechanical Engineering, Superfabulous University}

%----------------------------------FORMATTING-----------------------------------
\begin{document}
\maketitle
\linenumbers % start line numbers
\def\linenumberfont{\normalfont\small\rmfamily} % change line number font

%----------------------------------KEYWORDS-------------------------------------
\section*{Keywords}
Stuff, things, neat, cool, wow, instafun, tags4likes, etc

%----------------------------------ABSTRACT-------------------------------------
\section*{Abstract}
This is the text of the abstract.

%---------------------------------INTRODUCTION----------------------------------
\section*{Introduction}
For centuries, humankind has wondered: If I have two apples, and someone gives me another two apples, how many apples do I have? Some people did this \citep{Darwin1859}.

%-----------------------------------METHODS-------------------------------------
\section*{Methods}
We use the general framework outlined by Shelton et al (CITE).
That study outlined the structure for estimation of the proportional biomass of a taxon ($B_i$) given the proportional counts of sequences recovered from a parallel sequencing run ($Z_i$).
We constructed the following mathematical model (\ref{my_equation_label}) to better understand the concept:
% using '\ref{something}' in the text will refer to any object (e.g. figure, equation, table) which contains the corresponding '\label{something}'
% Note that you need to run latex a few times to get it to register numbers correctly

\begin{equation}\label{my_equation_label}
	Y = 2a + 2a
\end{equation}

Where $a$ represents an apple.

%-----------------------------------RESULTS-------------------------------------
\section*{Results}
We found that if you have two apples, and someone gives you another two apples, you have four apples.

%----------------------------------DISCUSSION-----------------------------------
\section*{Discussion}
Boy those results sure are neat. Now, the pressing question becomes: How do you like them apples?

%-------------------------------ACKNOWLEDGEMENTS--------------------------------
\section*{Acknowledgements}
We wish to thank all of the little people.

%-----------------------------------FUNDING-------------------------------------
\section*{Funding}
This study was funded by our super-rich uncle.

%--------------------------------CONTRIBUTIONS----------------------------------
\section*{Author Contributions}
Conceived and designed the experiments: .
Collected the data: .
Conducted the analyses: .
Wrote the first draft: .
Edited the manuscript: .

%-------------------------------------DATA-------------------------------------%
\section*{Data Availablity}
% Are the data and code available in a permanent, publicly accessible data archive or repository?
The data and code used to generate our results can be found at the following url:

%----------------------------------REFERENCES-----------------------------------
%\section*{References} % commented out because the section title is automatically inserted if using an automatically-generated bibliography

\bibliographystyle{apalike} % or: plain,unsrt,alpha,abbrv,acm,apalike,ieeetr
\bibliography{/Users/threeprime/Documents/Publications/bibtex/library} % path to your .bib file excluding .bib extension (e.g. /Users/threeprime/Documents/Publications/bibtex/library)

%-----------------------------------FIGURES-------------------------------------
\section*{Figures}

%\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
%  \centering
%    \includegraphics[width=1\textwidth]{figure_filename.pdf}
%    \caption{This is the figure caption.}
%  \label{myfigure} % use this to refer to your figure in the text, so that numbering updates automatically
%\end{figure}

%----------------------------------SUPPLEMENT-----------------------------------
%\section*{Supplemental Material}

\end{document}