### 2016-03-04 Manuscript/carkeek_grid.tex
add to introduction

%--------------------------------------||--------------------------------------%

\documentclass[11pt,letterpaper]{article} %  font size

%-----------------------------------PACKAGES-----------------------------------%

\usepackage[T1]{fontenc} % Choose an output font encoding (T1) that has support for the accented characters used by the most widespread European languages
\usepackage[utf8]{inputenc} % Allow input of accented characters (and more...)
\usepackage{graphicx} %  figures
\usepackage[round]{natbib} % names in citations
\usepackage{lineno} % line numbers
\usepackage{authblk} % allows more intuitive formatting for multiple authors/affiliations
\usepackage[margin=1in]{geometry} % make the margins 1 inch on all sides of the document
%\usepackage{amsmath} % useful for formatting math stuff, especially complex equations
%\usepackage{pdflscape} % rotate table into landscape mode
%\usepackage{subfigure} % side by side figures
%\usepackage{longtable} % for tables that span multiple pages.
\usepackage{setspace} % double spacing

%----------------------------------FORMATTING-----------------------------------

%\topmargin -1cm %0.0cm
%\textwidth 16cm % what does this do?
%\textheight 21cm % what does this do?
%\footskip 1.0cm % what does this do?
%\oddsidemargin 0.0cm

\date{\today}
\doublespacing % initiate double spacing (package setspace)
%\linespread{2} % alternate method of double spacing

%------------------------------------TITLE--------------------------------------

\title{There once was a grid at ol' Carkeek}

%-----------------------------------AUTHORS-------------------------------------

% using package 'authblk':
\author[1]{First Author\thanks{first.author@funstuff.com}}

\author[1,2]{Second Author}

\author[2]{Third Author}

\affil[1]{Department of Computer Science, \LaTeX\ University}
\affil[2]{Department of Mechanical Engineering, Superfabulous University}

%----------------------------------FORMATTING-----------------------------------
\begin{document}
\maketitle
\linenumbers % start line numbers
\def\linenumberfont{\normalfont\small\rmfamily} % change line number font

%----------------------------------KEYWORDS-------------------------------------
\section*{Keywords}
Stuff, things, neat, cool, wow, instafun, tags4likes, etc

%----------------------------------ABSTRACT-------------------------------------
\section*{Abstract}
This is the text of the abstract.

%---------------------------------INTRODUCTION----------------------------------
\section*{Introduction}

Biodiversity surveillance is being revolutionized by DNA-based detection of organisms from environmental samples. ?(specifically speed and scope of ecological studies).
Many researchers are justifiably cautious about the ?(adoption) of this new form of data.
Their apprehension is rooted in the premise that traditional survey approaches are more accurate because the chain of inference between observation and ecological data is usually short: A researcher sees two swans in Lake Hopatcong and infers the lake is occupied by at least 2 swans.
DNA based surveys, on the other hand, consist of a longer chain of inference ?(including several links at which the links are complex or unknown)?:
Dna sequences are observed from a sequencer, the sequencer sequences products of a polymerase chain reaction (PCR), PCR amplified sequences from a purified DNA sample, DNA is purified (extracted) from an environmental sample, environmental samples contain DNA from organisms present, the organisms present are representative of the biological community about which we wish to make inference. (Could tie this to the swans of Lake Hopatcong).
Clearly, this process is more complex than visual surveys. ?(and includes several steps at which the relationship between stages is complex or unknown)

%-----------------------------------METHODS-------------------------------------
\section*{Methods}
We use the general framework outlined by Shelton et al (CITE).
That study outlined the structure for estimation of the proportional biomass of a taxon ($B_i$) given the proportional counts of sequences recovered from a parallel sequencing run ($Z_i$).

We modeled the counts of DNA sequences ($Z$) from each of a given taxon $i$, in each replicate PCR $j$, from each replicate of a given location $k$ (hence, $Z_{ijk}$), as though they are ?(proportional to/drawn from)? a Poisson distribution. A Poisson distribution is described by one and only one parameter, $\lambda$, which is equal to both the mean and variance. Because in this case our modeled values are discrete counts, we use the natural exponent, $e^\lambda$. % WHY IS IT EXPONENTIATED? COUNTS?
Thus,
\begin{equation}\label{some_cool_eqn_name}
	Z_{ijk} \sim Poisson(e^{\lambda_{ijk}})
\end{equation}

In turn, we further assume this parameter $\lambda$ is linearly proportional to a suite of taxon-, pcr-, and site- specific parameters describing the variance associated with each sub-process linking the amount of DNA ($Y$) of a given taxon $i$ at a given location $k$ in a DNA extract (hence $Y_{ik}$):

\begin{equation}\label{GLM}
	\lambda_{ijk} = \beta_0 + \beta_i + \eta_{ijk} + \epsilon_{ijk}
\end{equation}

Where $\beta_0$ is a general intercept across all taxa, $\beta_i$ is a fixed effect accounting for the variance associated with taxon $i$, and $\eta_{ijk}$ and $\epsilon_{ijk}$ are random effects of variance resulting from the processes associated with PCR and spatial location, respectively.

%We constructed the following mathematical model (\ref{my_equation_label}) to better understand the concept:
% using '\ref{something}' in the text will refer to any object (e.g. figure, equation, table) which contains the corresponding '\label{something}'
% Note that you need to run latex a few times to get it to register numbers correctly

%\begin{equation}\label{my_equation_label}
%	Y = 2a + 2a
%\end{equation}
%
%Where $a$ represents an apple.

%-----------------------------------RESULTS-------------------------------------
\section*{Results}
We found that if you have two apples, and someone gives you another two apples, you have four apples.

%----------------------------------DISCUSSION-----------------------------------
\section*{Discussion}
Boy those results sure are neat. Now, the pressing question becomes: How do you like them apples?

%-------------------------------ACKNOWLEDGEMENTS--------------------------------
\section*{Acknowledgements}
We wish to thank all of the little people.

%-----------------------------------FUNDING-------------------------------------
\section*{Funding}
This study was funded by our super-rich uncle.

%--------------------------------CONTRIBUTIONS----------------------------------
\section*{Author Contributions}
Conceived and designed the experiments: James L. O'Donnell, Ryan P. Kelly, A. Ole Shelton.
Collected the data: James L. O'Donnell, Greg Williams, Natalie C. Lowell, Ryan P. Kelly, A. Ole Shelton, Jameal F. Samhouri.
Conducted the analyses: .
Wrote the first draft: .
Edited the manuscript: .

%-------------------------------------DATA-------------------------------------%
\section*{Data Availablity}
% Are the data and code available in a permanent, publicly accessible data archive or repository?
The data and code used to generate our results can be found at the following url:

%----------------------------------REFERENCES-----------------------------------
%\section*{References} % commented out because the section title is automatically inserted if using an automatically-generated bibliography

\bibliographystyle{apalike} % or: plain,unsrt,alpha,abbrv,acm,apalike,ieeetr
\bibliography{/Users/threeprime/Documents/Publications/bibtex/library} % path to your .bib file excluding .bib extension (e.g. /Users/threeprime/Documents/Publications/bibtex/library)

%-----------------------------------FIGURES-------------------------------------
\section*{Figures}

%\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
%  \centering
%    \includegraphics[width=1\textwidth]{figure_filename.pdf}
%    \caption{This is the figure caption.}
%  \label{myfigure} % use this to refer to your figure in the text, so that numbering updates automatically
%\end{figure}

%----------------------------------SUPPLEMENT-----------------------------------
%\section*{Supplemental Material}

\end{document}