### 2016-03-18 Manuscript/carkeek_grid.tex
add to manuscript

%--------------------------------------||--------------------------------------%

\documentclass[11pt,letterpaper]{article} %  font size

%-----------------------------------PACKAGES-----------------------------------%

\usepackage[T1]{fontenc} % Choose an output font encoding (T1) that has support for the accented characters used by the most widespread European languages
\usepackage[utf8]{inputenc} % Allow input of accented characters (and more...)
\usepackage{graphicx} %  figures
\usepackage[round]{natbib} % names in citations
\usepackage{lineno} % line numbers
\usepackage{authblk} % allows more intuitive formatting for multiple authors/affiliations
\usepackage[margin=1in]{geometry} % make the margins 1 inch on all sides of the document
%\usepackage{amsmath} % useful for formatting math stuff, especially complex equations
%\usepackage{pdflscape} % rotate table into landscape mode
%\usepackage{subfigure} % side by side figures
%\usepackage{longtable} % for tables that span multiple pages.
\usepackage{setspace} % double spacing

%----------------------------------FORMATTING-----------------------------------

%\topmargin -1cm %0.0cm
%\textwidth 16cm % what does this do?
%\textheight 21cm % what does this do?
%\footskip 1.0cm % what does this do?
%\oddsidemargin 0.0cm

\date{\today}
\doublespacing % initiate double spacing (package setspace)
%\linespread{2} % alternate method of double spacing

%------------------------------------TITLE--------------------------------------

\title{There once was a grid at ol' Carkeek}

%-----------------------------------AUTHORS-------------------------------------

% using package 'authblk':
\author[1]{First Author\thanks{first.author@funstuff.com}}

\author[1,2]{Second Author}

\author[2]{Third Author}

\affil[1]{Department of Computer Science, \LaTeX\ University}
\affil[2]{Department of Mechanical Engineering, Superfabulous University}

%----------------------------------FORMATTING-----------------------------------
\begin{document}
\maketitle
\linenumbers % start line numbers
\def\linenumberfont{\normalfont\small\rmfamily} % change line number font

%----------------------------------KEYWORDS-------------------------------------
\section*{Keywords}
Stuff, things, neat, cool, wow, instafun, tags4likes, etc

%----------------------------------ABSTRACT-------------------------------------
\section*{Abstract}
This is the text of the abstract.

%---------------------------------INTRODUCTION----------------------------------
\section*{Introduction}

Biodiversity surveillance is being revolutionized by DNA-based detection of organisms from environmental samples. ?(specifically speed and scope of ecological studies).
Many researchers are justifiably cautious about the ?(adoption) of this new form of data.
Their apprehension is rooted in the premise that traditional survey approaches are more accurate because the chain of inference between observation and ecological data is usually short: A researcher sees two swans in Lake Hopatcong and infers the lake is occupied by at least 2 swans.
DNA based surveys, on the other hand, consist of a longer chain of inference:
DNA sequences are reported by a sequencing machine, the machine identifies the sequence of products of a polymerase chain reaction (PCR), PCR amplifies pieces of DNA from a purified genomic DNA sample, DNA is purified (extracted) from an environmental sample, environmental samples contain DNA from organisms present, the organisms present are representative of the biological community about which we wish to make inference. ?( reverse order? tie to concrete example (swans of Lake Hopatcong)).
Clearly, this process is more complex than visual surveys, as the relationship between several steps is complex or unknown.
But consider that the processes ?(behind | underlying) other more widely-used ecological survey techniques are similarly complex, such as bird surveys based on song, or visual identification of fungal spores. When alternate survey approaches are impossible or inefficient, we are more willing to accept any available survey data, regardless of the complexity or uncertainty underlying it. (microbiologists have enthusiastically relied on DNA-based surveys for years for this reason, (though yes, they also do not have the problem of disconnect between individual and cell)).

The ability of DNA surveys to make quantitative inference about communities has been touted by some (CITE new fish quantitation paper) and doubted by others (CITE european eelgrass PLOSONE).
For example, a study linking (blah blah blah) concluded that "metabarcoding is powerful, yet blind" (CITE european eelgrass).
Conversely, others have reported strong quantitative and intuitive links between DNA-based and traditional survey methods (CITE Port 2016 MOLECO).
These studies usually rely on simple statistical models to link DNA quantity to some measurable ecosystem property like biomass (but see CITE).
When confronted with data collected in ?(complex ways/studies/whatever), simple models ?(may | often) fail to detect relationships when they exist, or vice versa ?(they are prone to inflated risk of BOTH type I and type II error) (CITE, see Woltman 2012).
For example, (CITE, look for that Gelman paper) have demonstrated that when data are structured in a hierarchical fashion (e.g. test scores of students in schools belonging to districts belonging to states), a low number of replicates at the first level of hierarchy (SEE THE PAPER).
Similarly, (describe hospital/school problems).

Shelton et al. (CITE Shelton 2016) outlined an approach for structuring statistical models of DNA surveys that address these issues.
This framework improved on alternative statistical techniques by explicitly accounting for the ?(hierarchical | nested | multilevel) structure of the study design, which allows error and uncertainty at each level to be ?(explicitly accounted for| modeled | propagated throughout the model).
That study demonstrated an improvement in the estimate of higher-level (e.g. ecological community) quantities when the processes linking them to the data are specified.
As an example, it was shown that incorporation of data about the mismatch between primer and template DNA sequence can improve the estimate of the relative abundance of unique DNA templates input to a PCR.

Here, we apply this framework to a DNA survey of ?(nearshore | coastal) marine habitat.
(TODO add commentary on current dogma surrounding distribution of DNA in well-mixed (marine) habitats).
We document the variability associated with lab based ?(procedures | replication | treatment; i.e. filter+DNA+PCR+seq), and the spatial scale over which DNA communities vary in this habitat.
We ?(show that | tested whether) a taxon's spatial distribution predicts ( the slope of the relationship between distance from shore and DNA abundance ~or~ to what degree DNA abundance is explained by distance from shore for each taxon).
We focus partly on species with known life histories that define their spatial distribution (e.g. shallow water livebearing fishes or sessile intertidal organisms with ?(motile/planktonic/pelagic) larvae or gametes).
For these taxa whose spatial distribution is well-documented and restricted, we calculate the rate of change in space and compare this rate among taxa with similar spatial distributions.
In turn, the distribution of rate of change serves as an estimate of the spatial distribution of DNA in this habitat.

We would love to estimate the minimum distance over which eDNA community differences can be detected.

%-----------------------------------METHODS-------------------------------------
\section*{Methods}
\subsection*{Environmental Sampling}
We collected samples from 8 points along three parallel transects separated by 1000 meters.
The first sample was collected over a lower-intertidal patch of \textit{Zostera marina}, with samples taken at 75, 125, 250, 500, 1000, 2000, and 4000 meters along the transect.

\subsection*{Laboratory Methods}

Samples were randomly assigned to PCR primer and library adapter index sequences. 
The sequencing run consisted of 14 samples ('libraries') prepared using different index sequences ligated during library preparation.
Of these libraries, ten comprised of amplicons prepared using the 16S protocol reported above, and four comprised of amplicons prepared using a 12S protocol similar to that reported by (CITE PORT 2015).

Pooled libraries were sequenced on an Illumina NextSeq at the Stanford Center for Functional Genomics (machine ID: NS50061; run ID: 115; flowcell ID: H3LFLAFXX).

\subsection*{Data Preparation (Bioinformatics)}
We calculated rates of cross-library contamination by counting occurrences of primer sequences: 12S primer sequences appearing in a 16S library (and vice versa) indicate an error in the preparation or sequencing procedures.

\subsection*{Community Analysis}
% Used k-medoids (PAM) rather than k-means because it does not require euclidean distances can be computed and that the observations follow a gaussian distribution
% because of the nature of our data (lots of zeros, with some very high counts), the mean is not very informative
We simultaneously assessed the existence of distinct community types and the membership of samples to those community types using a partitioning around mediods algorithm (CITE PAM, sometimes referred to as k-mediods clustering), as implemented in the R package fpc (CITE fpc).
The classification of samples to communities was made on the basis of their pairwise Bray-Curtis dissimilarity, calculated using the function vegdist in the R package vegan (CITE VEGAN). % consider Gower distance

% Calculate Moran's I on ?(relative abundance | raw count) of ?(taxa | OTU | uniq_sequences) to report dispersion of DNA in space?

\subsection*{Spatial Model Formulation}
We use the general framework outlined by Shelton et al (CITE).
That study outlined the structure for estimation of the proportional biomass of a taxon ($B_i$) given the proportional counts of sequences recovered from a parallel sequencing run ($Z_i$).

We modeled the counts of DNA sequences ($Z$) from each of a given taxon $i$, in each replicate PCR $j$, from each replicate of a given location $k$ (hence, $Z_{ijk}$), as though they are ?(proportional to/drawn from)? a Poisson distribution. A Poisson distribution is described by one and only one parameter, $\lambda$, which is equal to both the mean and variance. Because in this case our modeled values are discrete counts, we use the natural exponent, $e^\lambda$. % WHY IS IT EXPONENTIATED? COUNTS?
Thus,
\begin{equation}\label{some_cool_eqn_name}
	Z_{ijk} \sim Poisson(e^{\lambda_{ijk}})
\end{equation}

In turn, we further assume this parameter $\lambda$ is linearly proportional to a suite of taxon-, pcr-, and site- specific parameters describing the variance associated with each sub-process linking the amount of DNA ($Y$) of a given taxon $i$ at a given location $k$ in a DNA extract (hence $Y_{ik}$):

\begin{equation}\label{GLM}
	\lambda_{ijk} = \beta_0 + \beta_i + \eta_{ijk} + \epsilon_{ijk}
\end{equation}

%(\ref{my_equation_label})
Where $\beta_0$ is a general intercept across all taxa, $\beta_i$ is a fixed effect accounting for the variance associated with taxon $i$, and $\eta_{ijk}$ and $\epsilon_{ijk}$ are random effects of variance resulting from the processes associated with PCR and spatial location, respectively.

%We constructed the following mathematical model (\ref{my_equation_label}) to better understand the concept:
% using '\ref{something}' in the text will refer to any object (e.g. figure, equation, table) which contains the corresponding '\label{something}'
% Note that you need to run latex a few times to get it to register numbers correctly

%\begin{equation}\label{my_equation_label}
%	Y = 2a + 2a
%\end{equation}
%
%Where $a$ represents an apple.

%-----------------------------------RESULTS-------------------------------------
\section*{Results}
\subsection*{Data Quality (Bioinformatics)}
All value ranges are reported as (mean $\pm$ standard deviation).\\
There was a very low frequency of cross-contamination from other libraries into those reported here (5e-05$\pm$8e-05; max 0.00034)

\subsection*{Community Analysis}

\subsection*{Spatial Model Output}

%----------------------------------DISCUSSION-----------------------------------
\section*{Discussion}
Boy those results sure are neat. Now, the pressing question becomes: How do you like them apples?

%-------------------------------ACKNOWLEDGEMENTS--------------------------------
\section*{Acknowledgements}
We wish to thank all of the little people.

%-----------------------------------FUNDING-------------------------------------
\section*{Funding}
This study was funded by our super-rich uncle.

%--------------------------------CONTRIBUTIONS----------------------------------
\section*{Author Contributions}
Conceived and designed the experiments: James L. O'Donnell, Ryan P. Kelly, A. Ole Shelton.
Collected the data: James L. O'Donnell, Greg Williams, Natalie C. Lowell, Ryan P. Kelly, A. Ole Shelton, Jameal F. Samhouri.
Conducted the analyses: .
Wrote the first draft: .
Edited the manuscript: .

%-------------------------------------DATA-------------------------------------%
\section*{Data Availablity}
% Are the data and code available in a permanent, publicly accessible data archive or repository?
The data and code used to generate our results can be found at the following url:

%----------------------------------REFERENCES-----------------------------------
%\section*{References} % commented out because the section title is automatically inserted if using an automatically-generated bibliography

\bibliographystyle{apalike} % or: plain,unsrt,alpha,abbrv,acm,apalike,ieeetr
\bibliography{/Users/threeprime/Documents/Publications/bibtex/library} % path to your .bib file excluding .bib extension (e.g. /Users/threeprime/Documents/Publications/bibtex/library)

%-----------------------------------FIGURES-------------------------------------
\section*{Figures}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{../Figures/site_map.pdf}
    \caption{Geographic position of collected samples. Lines give XXX meter isobaths.}
  \label{site_map} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{clustering_fig.pdf}
    \caption{Plot of ?(non-metric multi-dimensional scaling | principal components ) analysis. Points are colored by their membership to clusters of k-means clustering analysis.}
  \label{clustering_fig} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{dissim_by_distance.pdf}
    \caption{Pairwise dissimilarity ?(BRAY CURTIS/ETC) of eDNA communities plotted against pairwise spatial distance.}
  \label{dissim_by_distance} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{slope_plots.pdf}
    \caption{Fit lines of DNA sequence counts as a function of distance from shore for a selection of taxa for which we have strong preconceived expectations (left). Box plots of the estimates of the slopes for taxa ?(100 most abundant)?, grouped by life history traits (right).}
  \label{slope_plots} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{var_boxplots.pdf}
    \caption{Box plots of estimates of variance associated with each level of the multilevel model, corresponding to stages of the eDNA sampling protocol.}
  \label{variance_boxplots} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{clustering_map.pdf}
    \caption{Geographic position of collected samples, colored by membership to clusters identified by k-means clustering analysis. Lines give 10m isobaths.}
  \label{clustering_map} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

%\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
%  \centering
%    \includegraphics[width=1\textwidth]{figure_filename.pdf}
%    \caption{This is the figure caption.}
%  \label{myfigure} % use this to refer to your figure in the text, so that numbering updates automatically
%\end{figure}

%----------------------------------SUPPLEMENT-----------------------------------
%\section*{Supplemental Material}

\end{document}