### 2016-08-13 Manuscript/v0.0/carkeek_grid.tex
removed irrelevant text

%--------------------------------------||--------------------------------------%

\documentclass[11pt,letterpaper]{article} %  font size

%-----------------------------------PACKAGES-----------------------------------%

\usepackage[T1]{fontenc} % Choose an output font encoding (T1) that has support for the accented characters used by the most widespread European languages
\usepackage[utf8]{inputenc} % Allow input of accented characters (and more...)
\usepackage{textcomp,marvosym} % additional figures, including textdegree
\usepackage{graphicx} %  figures
\usepackage[round]{natbib} % names in citations
\usepackage{lineno} % line numbers
\usepackage{authblk} % allows more intuitive formatting for multiple authors/affiliations
\usepackage[margin=1in]{geometry} % make the margins 1 inch on all sides of the document
%\usepackage{amsmath} % useful for formatting math stuff, especially complex equations
%\usepackage{pdflscape} % rotate table into landscape mode
%\usepackage{subfigure} % side by side figures
%\usepackage{longtable} % for tables that span multiple pages.
\usepackage{setspace} % double spacing

%----------------------------------FORMATTING-----------------------------------

%\topmargin -1cm %0.0cm
%\textwidth 16cm % what does this do?
%\textheight 21cm % what does this do?
%\footskip 1.0cm % what does this do?
%\oddsidemargin 0.0cm

\date{\today}
\doublespacing % initiate double spacing (package setspace)
%\linespread{2} % alternate method of double spacing
%\graphicspath{{../../Figures/}} % I think if you do this it saves typing the path over and over in the figures section

%-------------------------------------------------------------------------------

\title{There once was a grid at ol' Carkeek}
% Spatial dispersion of environmental DNA in a marine habitat

%-----------------------------------AUTHORS-------------------------------------

% using package 'authblk':
\author[1]{First Author\thanks{first.author@funstuff.com}}

\author[1,2]{Second Author}

\author[2]{Third Author}

\affil[1]{Department of Computer Science, \LaTeX\ University}
\affil[2]{Department of Mechanical Engineering, Superfabulous University}

%----------------------------------FORMATTING-----------------------------------
\begin{document}
\maketitle
\linenumbers % start line numbers
\def\linenumberfont{\normalfont\small\rmfamily} % change line number font

%----------------------------------KEYWORDS-------------------------------------
\section*{Keywords}
Stuff, things, neat, cool, wow, instafun, tags4likes, etc

%-------------------------------------------------------------------------------
%----------------------------------ABSTRACT-------------------------------------
%-------------------------------------------------------------------------------
\section*{Abstract}
This is the text of the abstract.

%-------------------------------------------------------------------------------
%---------------------------------INTRODUCTION----------------------------------
%-------------------------------------------------------------------------------
\section*{Introduction}

Samples collected of ecological communities may vary in dissimilarity from 0 (completely identical) to 1 (completely different).
For samples collected from multiple locations, the relationship between their spatial distance and community dissimilarity is of interest because it reflects the amount of community heterogeneity---changes in abundance and composition---over the spatial scale sampled.
%In general, the relationship between community dissimilarity (0 = identical; 1 = completely different) and spatial distance is expected to be asymptotic, because communities nearer to each other tend to be more similar than those farther apart.
The intercept is expected to be 0, because only within-sample comparisons can have 0 spatial separation, and communities have no dissimilarity within a sample if sampling method is repeatable. % technically impossible for samples to be taken at exactly 0; they are actually immediately adjacent.
%Deviation from 0 indicates heterogeneous community composition/structure over fine scales.
Likewise, dissimilarity cannot exceed 1, and reaches 1 only when multiple discrete community types are sampled.
The trend is expected to be asymptotic if communities within a habitat are spatially structured, where the value of the asymptote and the rate of increase provide insight about community turnover.
A flat relationship indicates that heterogeneity is not assorted spatially, and can be interpreted in different ways, depending on the asymptote. % ?or no relationship both
If the asymptote is close to 1, there is high spatial heterogeneity over the spatial scale of sampling.
If the asymptote is close 0, all samples are similar, and we infer there is complete community homogeneity over the scale sampled.
The rate at which community dissimilarity approaches the mean gives an indication of the rate of community turnover.

%-------------------------------------------------------------------------------
%-----------------------------------METHODS-------------------------------------
%-------------------------------------------------------------------------------
\section*{Methods}
\subsection*{Environmental Sampling}
Starting from lower-intertidal patches of \textit{Zostera marina}, we collected water samples at 1 meter depth from 8 points (0, 75, 125, 250, 500, 1000, 2000, and 4000 meters) along three parallel transects separated by 1000 meters (Figure \ref{site_map}).

\subsection*{Laboratory Procedures}

Samples were randomly assigned to PCR primer and library adapter index sequences.
The sequencing run consisted of 14 samples ('libraries') prepared using different index sequences ligated during library preparation.
Of these libraries, ten comprised of amplicons prepared using the 16S protocol reported above, and four comprised of amplicons prepared using a 12S protocol similar to that reported by (CITE PORT 2015).

Pooled libraries were sequenced on the Illumina NextSeq platform at the Stanford Center for Functional Genomics (machine ID: NS50061; run ID: 115; flowcell ID: H3LFLAFXX). Raw sequence data in fastq format is publicly available (see Data Availability).

\subsection*{Data Preparation (Bioinformatics)}

Detailed bioinformatic methods are provided in the supplemental material, and scripts used from raw sequencer output onward can be found in the project directory on GitHub (see Data Availability).

%----------------------------------------------------------------------------
%METHODS
%Sequences passing were demultiplexed on the basis of the ?(8-bp) sequence ligated during library preparation and read during the index read, and filtered for general quality by the sequencing facility.
%
%We used fastqc to assess the fastq files output from the sequencer for low-quality indications of a problematic run.
%
%Reads that contain primer index sequences (or combinations thereof) that we did not actually use in laboratory procedures are known to ?(be derived from contamination) be a result of processes that leave the same signature as contamination. For example, errors during oligonucleotide synthesis of the indexed primers would cause reads to be erroneously assigned to samples.
%Because this represents the minimum rate at which we can be certain that sequences from one environmental sample could be erroneously assigned, we considered for further analysis only those reads occurring with greater frequency than this.
%
We calculated rates of cross-library contamination by counting occurrences of primer sequences: 12S primer sequences appearing in a 16S library (and vice versa) indicate an error in the preparation or sequencing procedures.

We checked for experimental error by evaluating the Bray-Curtis dissimilarity of proportional read abundance among replicate PCRs of the same DNA sample (0.033$\pm$0.063), and excluded one PCR replicate for which the dissimilarities between itself and the other replicates exceeded 1 SD (lib\_B\_tag\_GCGCTC).

%To infer a general estimate of the rate of cross-sample contamination likely to be associated with lab procedures, we calculated the proportion of reads containing primer index sequences (or combinations thereof?) that we did not actually use
%
%RESULTS
%Reads were of generally high-quality, with ?(96.48)? % of base calls having a 0.001 or lower probability of incorrect base call (Phred q-score of 30 or higher).
%
%
%Total sequencing clusters passing filter per read: 371,576,190 on NEXTSEQ
%
% THESE BELOW ARE FROM A MISEQ RUN
%reads, reads passing filter, % >= Q30, for each run ( 7,052,307 ; 6,752,829; 97.47 :: 7,052,307 ; 6,752,829 ; 95.50)
%
%Cluster density of 358 \UTF{00EF}\UTF{00BF}\UTF{0153}22	clusters per mm^2
%
%XXX % of reads aligned to the PhiX genome
%(31.02 \UTF{00EF}\UTF{00BF}\UTF{0153}0.59	; 29.17 \UTF{00EF}\UTF{00BF}\UTF{0153}0.89)
%
%and error rate as determined from alignment of PhiX reads to the PhiX genome was low (Read 1: 1.26 \UTF{00EF}\UTF{00BF}\UTF{0153}0.07; Read 2: 1.38 \UTF{00EF}\UTF{00BF}\UTF{0153}0.07)
%----------------------------------------------------------------------------

To account for variation in the number of sequencing reads (sequencing depth) recovered per sample, we multiplied the within-sample proportional abundance of each OTU by the minimum sequencing depth (130402), and rounded to the nearest integer.

Because each step in the massively parallel sequencing workflow is sensitive to contamination, it is possible that some sequences are the result of contamination during field sampling, filtration, DNA extraction, PCR, fragment size selection, quantitation, sequencing adapter ligation, or the sequencing process itself.
Some authors have argued that these risks could bias sequence abundance, making those data meaningless and prohibiting quantitative estimates, yet convert count data to binary presence absence data on the basis of the sequence abundance greater than some arbitrary threshold). %massively parallel metabarcoding,
Recent work has shown that this binary treatment of data can (?falsely?) overestimate taxon richness and falsely elevate the estimate of taxon turnover among samples (CITE LERAY FORTHCOMING).
We take the view that contaminants are unlikely to manifest as sequences in the final dataset in consistent abundance across replicates; indeed, our data show that the process from PCR onward is remarkably consistent.
Thus, we calculated from our data the maximum number of sequence counts (after scaling to correct for sequencing depth variation) for which there is turnover in presence-absence among PCR replicates within an environmental sample.
We use this number to determine a conservative minimum threshold above which we can be confident that counts are consistent among replicates and not of (?spurious | dubious?) origin, and exclude from further analysis observations where the mean abundance across PCR replicates within samples does not reach this threshold. % the number is turnover_threshold=178

\subsection*{Ecological Analyses}

We subset the data in a variety of ways and conducted each analysis on all subsets.
We report the subset used with each analysis, and report results on alternative subsets in the supplemental material.
For all analyses beyond the assessment of PCR consistency, we use the mean taxon abundance across PCR replicates from each of the 24 environmental samples.
Our subsetting methods were (1) exclude rare taxa ?(threshold)?, (2) exclude abundant taxa ?(threshold)?, (3) subsampling of taxa randomly, (4) subsampling of taxa proportional to their abundance, (5) subsampling of taxa inversely proportional to their abundance, (6) exclude taxa found in only one environmental sample (spatially invariant), (7) exclude non-marine taxa (e.g. humans, pigs), (8) exclude taxa whose known individual range (including gametes and larvae) exceeds the spatial scale of our study.
We also tested a variety of transformations of the mean scaled abundance data, including (1) log ($log_{e}x$), and (2) binary ($1 = x > 1$; $0 = x < 1$).

% Used k-medoids (PAM) rather than k-means because it does not require euclidean distances can be computed and that the observations follow a gaussian distribution
% because of the nature of our data (lots of zeros, with some very high counts), the mean is not very informative
We simultaneously assessed the existence of distinct community types and the membership of samples to those community types using a partitioning around mediods algorithm (CITE PAM, sometimes referred to as k-medoids clustering), as implemented in the R package fpc (CITE fpc).
The classification of samples to communities was made on the basis of their pairwise Bray-Curtis dissimilarity, calculated using the function vegdist in the R package vegan (CITE VEGAN). % consider Gower distance

We calculated the great circle distance between points using the Haversine method as implemented by the R package geosphere (CITE geosphere).
% Calculate Moran's I on ?(relative abundance | raw count) of ?(taxa | OTU | uniq_sequences) to report dispersion of DNA in space?

To estimate the maximum dissimilarity and the rate of community turnover in space, we modeled community dissimilarity as a function of distance from shore following a Michaelis-Menten relationship:
\begin{equation}\label{MichaelisMenten}
	com \sim V_{max}[d]/K_{m}[d]
\end{equation}
where $com$ is community dissimilarity, $d$ is spatial distance, and where the asymptote is given by $V_{max}$, and the distance at which half the asymptote has been reached is given by $K_{m}$.
Model fit was assessed using the function nls in R (CITE R).

\subsection*{Organismal Life History Data}
We collated coarse-scale data on life history characteristic for each of the major taxonomic groups recovered,
including dispersal range of the gametes, larvae, and adults, adult habitat type and selectivity, and adult body size.
Dispersal range was given as an order-of-magnitude approximation of the scale of dispersal: for example, internally fertilized species were assigned a gamete range of 0 km, while broadcast spawners were assigned a gamete range of 10 km.
Similarly, adult range size was approximated as 0 km (sessile), 1 km (motile but not pelagic), or 10 km (highly mobile, pelagic).
Variables were specified as 'multiple' for groups known to span more than 1 magnitude of range size.
For groups to which sequences were annotated with high confidence, but for which life history strategy is diverse or poorly known (e.g. families in the phylum Nemertea), we used conservative, coarse approximations at a higher taxonomic rank.
%We assessed whether or not marine invertebrate taxa are thought to be present in Puget Sound by checking for their presence in a comprehensive checklist of invertebrates of Puget Sound CITE KOZLOFF.
These data are available as part of the REFERENCE SUPPLEMENTAL DATA.

\subsection*{Spatial Model Formulation}
We use the general framework outlined by Shelton et al (CITE).
That study outlined the structure for estimation of the proportional biomass of a taxon ($B_i$) given the proportional counts of sequences recovered from a parallel sequencing run ($Z_i$).

We modeled the counts of DNA sequences ($Z$) from each of a given taxon $i$, in each replicate PCR $j$, from each replicate of a given location $k$ (hence, $Z_{ijk}$), as though they are ?(proportional to/drawn from)? a Poisson distribution. A Poisson distribution is described by one and only one parameter, $\lambda$, which is equal to both the mean and variance. Because in this case our modeled values are discrete counts, we use the natural exponent, $e^\lambda$. % WHY IS IT EXPONENTIATED? COUNTS?
Thus,
\begin{equation}\label{some_cool_eqn_name}
	Z_{ijk} \sim Poisson(e^{\lambda_{ijk}})
\end{equation}

In turn, we further assume this parameter $\lambda$ is linearly proportional to a suite of taxon-, pcr-, and site- specific parameters describing the variance associated with each sub-process linking the amount of DNA ($Y$) of a given taxon $i$ at a given location $k$ in a DNA extract (hence $Y_{ik}$):

\begin{equation}\label{GLM}
	\lambda_{ijk} = \beta_0 + \beta_i + \eta_{ijk} + \epsilon_{ijk}
\end{equation}

%(\ref{my_equation_label})
Where $\beta_0$ is a general intercept across all taxa, $\beta_i$ is a fixed effect accounting for the variance associated with taxon $i$, and $\eta_{ijk}$ and $\epsilon_{ijk}$ are random effects of variance resulting from the processes associated with PCR and spatial location, respectively.

%We constructed the following mathematical model (\ref{my_equation_label}) to better understand the concept:
% using '\ref{something}' in the text will refer to any object (e.g. figure, equation, table) which contains the corresponding '\label{something}'
% Note that you need to run latex a few times to get it to register numbers correctly

%\begin{equation}\label{my_equation_label}
%	Y = 2a + 2a
%\end{equation}
%
%Where $a$ represents an apple.

%-------------------------------------------------------------------------------
%-----------------------------------RESULTS-------------------------------------
%-------------------------------------------------------------------------------
\section*{Results}
\subsection*{Data Quality (Bioinformatics)}
All value ranges are reported as (mean $\pm$ standard deviation).\\
There was a very low frequency of cross-contamination from other libraries into those reported here (5e-05$\pm$8e-05; max 0.00034)

We assessed the consistency of PCR by conducting 4 replicate PCRs for each environmental sample and calculating the mean pairwise Bray-Curtis dissimilarity of the resulting communities (scaled to minimum read depth per sample).
92 of the 96 amplicon samples had mean Bray-Curtis dissimilarity $\leq$ 0.052; 1 sample had a value of 0.341, which elevates the value of the other replicates.
After removal of this sample, the highest mean Bray-Curtis dissimilarity among replicates within an environmental sample was 0.034.

\subsection*{Community Analysis}
Excluding spatially-invariant taxa ?(taxa which occur in only one spatial location) had no discernible effect on the outcome of the PAM analysis (number of clusters, assignment to clusters).

The estimated asymptote of community dissimilarity as a function of spatial distance ($V_m$) was 0.72 ($p <<< 0.05$), and the distance at which half this dissimilarity was accumulated ($K_m$) is 23.8 kilometers ($p = 0.006$).
Residual standard error of the fit of the model is 0.1563 on 274 degrees of freedom.

The vast majority (97.6\%) of sequences and OTUs (96.9\%) were matched to organisms that have high potential for dispersal at either the gamete, larval, or adult stage.
Of the 6 OTUs with limited dispersal, only 2 occurred in more than two samples.
These were assigned to \textit{Cymatogaster aggregata}, a viviparous nearshore fish with internal fertilization, and \textit{Cupolaconcha meroclista}, a sessile marine gastropod with internal fertilization and short larval dispersal (CITE???).
Thus, we are unable to compare the patterns of dispersion for varying life history categories.

\subsection*{Spatial Model Output}

%-------------------------------------------------------------------------------
%----------------------------------DISCUSSION-----------------------------------
%-------------------------------------------------------------------------------
\section*{Discussion}
Boy those results sure are neat. Now, the pressing question becomes: How do you like them apples?

%-------------------------------ACKNOWLEDGEMENTS--------------------------------
\section*{Acknowledgements}
We wish to thank all of the little people.

%-----------------------------------FUNDING-------------------------------------
\section*{Funding}
This study was funded by our super-rich uncle.

%--------------------------------CONTRIBUTIONS----------------------------------
\section*{Author Contributions}
Conceived and designed the experiments: James L. O'Donnell, Ryan P. Kelly, A. Ole Shelton.
Collected the data: James L. O'Donnell, Greg Williams, Natalie C. Lowell, Ryan P. Kelly, A. Ole Shelton, Jameal F. Samhouri.
Conducted the analyses: .
Wrote the first draft: .
Edited the manuscript: .

%-------------------------------------DATA-------------------------------------%
\section*{Data Availablity}
% TODO SUBMIT DATA AND METADATA; ADD URL
All sequence files and metadata are available from EMBL:\\ \verb!http://www.ebi.ac.uk/ena/data/view/XXXXXXXXX! \\
All analyses were performed using scripts available from the project repository on GitHub:\\  \verb!https://github.com/jimmyodonnell/Carkeek_eDNA_grid!\\

%----------------------------------REFERENCES-----------------------------------
%\section*{References} % commented out because the section title is automatically inserted if using an automatically-generated bibliography
\bibliographystyle{apalike} % or: plain,unsrt,alpha,abbrv,acm,apalike,ieeetr
\bibliography{carkeek_grid_bib} % path to your .bib file excluding .bib extension (e.g. /Users/threeprime/Documents/Publications/bibtex/library)

% 1. Jane SF, Wilcox TM, Mckelvey KS, Young MK, Schwartz MK, Lowe WH, et al. Distance, flow and PCR inhibition: EDNA dynamics in two headwater streams. Mol Ecol Resour. 2014;
%-----------------------------------FIGURES-------------------------------------
\section*{Figures}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{../../Figures/site_map.pdf}
    \caption{TODO: Plot with GEBCO 30-second data or remove grid coloring and color by isobath. Looking into filling by contour. Geographic position of collected samples. Lines give XXX meter isobaths.}
  \label{site_map} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{../../Figures/diss_by_dist.pdf}
    \caption{Pairwise Bray-Curtis dissimilarity of eDNA communities plotted against pairwise spatial distance.
    Line represents prediction of the Non-linear Least Squares regression to a Michaelis-Menten model ($V_m = 0.72$, $p <<< 0.05$; $K_m = 23.8$ kilometers, $p = 0.006$; RSE = 0.1563; df = 274.).
    Restricting comparison to within-transect has no qualitative difference in the outcome (see 'diss\_by\_dist\_by\_transect.pdf').
    }
  \label{comm_diss_by_geo_dist} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{../../Figures/dissimilarity_from_shore.pdf}
    \caption{Pairwise dissimilarity (Bray-Curtis) across transects plotted against distance from shore. A linear model determined no effect of distance from shore on dissimilarity.}
  \label{dissimilarity_from_shore} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{../../Figures/variance_from_shore.pdf}
    \caption{Within-OTU variance across transects plotted against distance from shore.}
  \label{variance_from_shore} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{../../Figures/heatmap_sites.pdf}
    \caption{Heatmap of pairwise site dissimilarity (Bray-Curtis).}
  \label{heatmap} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[height=0.5\textheight]{../../Figures/pam_plot.pdf}
    \caption{TODO figure out color of ellipses; I can't even plot them gray without Plot of partitioning around medoids (PAM) analysis of OTU sequence abundance from 4 replicate PCRs at each of 24 sampling points. Points represent communities of OTUs; color and shape indicate cluster membership as determined by PAM analysis. Ellipses indicate the smallest area of a cluster that contains all of its members.}
  \label{pam_plot.pdf} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[height=0.5\textheight]{../../Figures/pam_in_space.pdf}
    \caption{Geographic position of collected samples, colored by membership to clusters identified by partitioning around medoids algorithm.
%    Points are jittered in both horizontal and vertical dimension to distinguish among four replicate PCR products sequenced from each environmental sample.
}
  \label{clustering_map} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

%\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
%  \centering
%    \includegraphics[width=1\textwidth]{../../Figures/slope_plots.pdf}
%    \caption{Fit lines of DNA sequence counts as a function of distance from shore for a selection of taxa for which we have strong preconceived expectations (left). Box plots of the estimates of the slopes for taxa ?(100 most abundant)?, grouped by life history traits (right).}
%  \label{slope_plots} % use this to refer to your figure in the text, so that numbering updates automatically
%\end{figure}

%\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
%  \centering
%    \includegraphics[width=1\textwidth]{../../Figures/var_boxplots.pdf}
%    \caption{Box plots of estimates of variance associated with each level of the multilevel model, corresponding to stages of the eDNA sampling protocol.}
%  \label{variance_boxplots} % use this to refer to your figure in the text, so that numbering updates automatically
%\end{figure}

\begin{figure}[h!] % [h!] forces the figure to be placed roughly here
  \centering
    \includegraphics[width=1\textwidth]{../../Figures/otu_in_space.pdf}
    \caption{Example of a DNA sequence's spatial distribution. This sequence is annotated to SPECIES X, which is found only in shallow, structured habitats such as patches of \textit{Zostera marina}. Point size and color transparency indicates abundance relative to other DNA sequences from that sample, scaled to the maximum value for this sequence (no fill = 0, full fill = 1). Samples from which this sequence was not recovered are indicated by an "x".}
  \label{otu_in_space} % use this to refer to your figure in the text, so that numbering updates automatically
\end{figure}

%----------------------------------SUPPLEMENT-----------------------------------
\section*{Supplemental Material}

\subsection*{Bioinformatic Methods}

\end{document}